---
title: "ML-assignment"
author: "Manzar"
date: "10/18/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Practical Machine Learning
<br/>

first load the required libraries. 
```{r}
#free up some memory
rm(list = ls())

#libraries

library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(12345)
```


set the url for the download


```{r url for download}
urltrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urltest <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

#download

training <- read.csv(url(urltrain))
testing  <- read.csv(url(urltest))

#create partition for data validation

inTrain <- createDataPartition(training$classe, p=.7, list = FALSE)
trainset <- training[inTrain,]
testset <- training[-inTrain,]

dim(trainset); dim(testset)
```

Data cleaning. Remove variables/columns with near zero values

```{r }
nzv <- nearZeroVar(trainset)
trainset <- trainset[, -nzv]
testset <- testset[, -nzv]
```

Data cleaning. Remove variables which are mostly NA
```{r}
varNA <- sapply(trainset, function(x) mean(is.na(x))) > .95
trainset <- trainset[, varNA==F]
testset <- testset[, varNA==F]

```

Remove personal identifiers

```{r}
trainset <- trainset[, -(1:5)]
testset  <- testset[, -(1:5)]
```

## Correlation Analysis

```{r}
corMatrix <- cor(trainset[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", tl.cex = .6, 
         tl.col = rgb(0,0,0))
```

## Prediction Model Building

A) Method: Random Forest

```{r}

#creating model fit

set.seed(12345)
controlRF <- trainControl(method = "cv", number = 3, verboseIter = FALSE)
modFitRandForest <- train(classe~., data = trainset, method = "rf", trControl=controlRF)
modFitRandForest$finalModel
```


```{r}
#prediction on Test dataset

predictRandomForest <- predict(modFitRandForest, newdata = testset)
confMatRandForest <- confusionMatrix(predictRandomForest, as.factor(testset$classe))
confMatRandForest
```
```{r}
#plot matrix results

plot(confMatRandForest$table, col = confMatRandForest$byClass, main = paste("Random Forest - Accuracy =", 
                                round(confMatRandForest$overall['Accuracy'], 4)))

#table(factor(predictRandomForest, levels=min(testset):max(testset)), 
  #    factor(testset$classe, levels=min(testset):max(testset)))
```


## Method: Decision Trees

```{r}
#model fit

set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data = trainset, method = "class")
fancyRpartPlot(modFitDecTree)
```

```{r}
#prediction of Test dataset

predictDecTree <- predict(modFitDecTree, newdata = testset, type = "class")
confMatDecTree <- confusionMatrix(predictDecTree, as.factor(testset$classe))
confMatDecTree

#plot matrix result

plot(confMatDecTree$table, col = confMatDecTree$byClass, main = paste("Decision Tree - Accuracy =",
                               round(confMatDecTree$overall['Accuracy'], 4)))
```

## Generalized Boosted Model

```{r}
# model fit

set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=trainset, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```
```{r}
# prediction on Test dataset
predictGBM <- predict(modFitGBM, newdata=testset)
confMatGBM <- confusionMatrix(predictGBM, as.factor(testset$classe))
confMatGBM
```

```{r}
# plot matrix results
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))
```

## Applying the Selected Model to the Test Data

The accuracy of the 3 regression modeling methods above are:

Random Forest : 0.9963
Decision Tree : 0.7368
GBM           : 0.9839


In that case, the Random Forest model will be applied to predict the 
20 quiz results (testing dataset) as shown below.


```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```















